BOLD Parameter Estimation Using Particle Filters

### 2010-08-07 doc/thesis/conclusion.tex
more work with the conclusion, just need to do future works

% Enhancements/Dehancements 
%% flatten prior
%% \% difference using spline?
%% "smart knots"
%% DC gain as a parameter
%% linearizing

% Future works
%% Extensive test of quality of volterra kernel estimate from friston
%% Better de-trending
%% Automatic estimation of measurement error.

>>>>

% Enhancements/Dehancements 
%% flatten prior
%% \% difference using spline?
%% "smart knots"
%% DC gain as a parameter
%% linearizing
\section{Particle Filter Variations}
There are quite a few areas where the particle filter could be tuned. While
some of these areas were touched on \autoref{sec:Methods}, certain parts were
not used in the analysis and so weren't mentioned.

\subsection{Prior Distribution}
Starting with a flattened
prior (by setting weights to be non-uniform after the initial particles have
been drawn) was not used in the final analysis. The reason for not flattening
the prior was a practical issue with weighting points in a 
7 dimensional joint distribution. Often differences in particle weights approached machine 
precision, meaning that the flattened prior was actually far from flat. Instead
the distribution became unpredictable, which made the results unpredictable. This
is quantization issue that will exist unless the prior were made to be deterministic
or the initial number particles was raised to computationally prohibitive levels.

Wider prior

\subsection{Detrending}
A large number of de-trending methods were tested, but all had some major problem. 
One viable method that could be utilized given the right experimental design is detrending
based on areas of low activity. This has the advantage that it wouldn't require an arbitrary
constant to be added to the pre-processed signal for the BOLD model to fit properly.
The disadvantage of this approach is that it could hide long fall times by normalizing them
out. It also requires periodic breaks in the stimulus, which could reduce value of those
samples for fitting purposes.

Another potential way to deal with detrending is to 
linearize. This would use the delta between measurements for fitting rather than the 
direct value. This has the advantage of not requiring detrending and thus 
gives nearly raw data to the particle filter
for processing. The effectiveness of this method depends directly on the type of stimulus.
In an event driven stimulus with sparse stimuli this could be extremely effective because
the DC level has minimal data; however the longer high levels are held the less effective
this will be. On the other hand, splines will also reduce the plateaus, so its possible
linearizing may be just as effective. I found in tests that large drift-low white noise
type signals performed much better with a linearization approach, as one might expect. 
The results were equal to or worse in real data; but that was only for one particular
stimulus sequence.

A subtle difference that could be made in detrending is how the percent difference
is calculated. In this work the spline was subtracted, and then the result
was divided by the average of the initial signal. A more correct method may be dividing
by the spline value at that particular point. The reason for not dividing by the spline
at that point is that it could be less stable, and in a sense the trend has already
been removed. The difference is not particularly large though 
(dividing by 1020 rather than 1000 for instance) as long as the spline didn't have heavy swings.

Finally, rather than adding a constant to the detrended data before
applying the particle filter, a DC gain parameter could be added to the model. In
tests, this could work well, however, it often did not. The problem with this
could simply be the addition of another degree of freedom without any increase in
the number of particles. Increasing the number of particles further though
can be computationally intractable. Thus, while this is in a sense the correct
solution, it is an impractical one.

% Future works
%% Extensive test of quality of volterra kernel estimate from friston
%% Better de-trending
%% Automatic estimation of measurement error.
%todo
